Decoding UK Hospital Wait Times: Infrastructure, Workforce, and Demand

### 2024-11-26 paper/paper.qmd
Wrote datasheet for RTT data.

### Relationship between Wait Times and Type 1 Major Attendance

The weak positive association between attendance rates and wait times, though not significant, aligns with broader concerns about rising patient loads straining healthcare systems. As attendance increases, hospitals may prioritize emergency cases, diverting resources from elective and non-urgent procedures, potentially exacerbating delays. Factors such as limited access to primary care, changes in population demographics, and healthcare-seeking behaviors likely drive the observed trends. Although the relationship was not statistically significant, addressing the root causes of rising attendance remains critical to reducing systemic strain. Potential strategies include expanding primary care access, implementing community triaging systems, and improving public health education to promote appropriate healthcare utilization. These measures could help balance patient loads and reduce delays for non-emergency procedures.

@@

## Future Work
To address the limitations of this study, future research should prioritize collecting monthly data for the years 2015 to 2019 to explore more granular trends before the anomalies introduced by COVID-19. Analyzing these pre-pandemic patterns could provide a clearer understanding of the role that metrics such as beds per 1,000, physician density, and attendance rates played during a relatively stable period. Additionally, expanding the dataset to include broader healthcare metrics, such as regional funding levels, technological investments, and patient demographics, could yield a more comprehensive analysis. Future studies should also explore non-linear models or interaction effects to capture the intricate relationships between predictors. Simulation models and scenario analyses could be used to test policy interventions, such as optimizing resource allocation or improving workflow efficiency, to identify actionable strategies for reducing wait times and improving healthcare delivery.

@@

# Print the combined plot
print(combined_plot)
```
### Adapted Model 
```{r}
#| echo: false
#| output: asis
#| eval: true
#| label: tbl-modeladj
#| tbl-cap: "Linear model explaining percentage increase in wait times based on healthcare system predictors"

>>>>

### Relationship between Wait Times and Type 1 Major Attendance
The weak positive association between attendance rates and wait times, though not significant, aligns with broader concerns about rising patient loads straining healthcare systems. The relationship is likely weak due to the significant deviation from pre-Covid trend in 2020, with a sudden drop by 25% from 2019 levels. This drop was recovered in 2021, but the impact on the linear model constructed by a limited 8 observation data set cannot be ignored. As attendance increases, hospitals may prioritize emergency cases, diverting resources from elective and non-urgent procedures, potentially exacerbating delays. Factors such as limited access to primary care, changes in population demographics, and healthcare-seeking behaviors likely drive the observed trends. Although the relationship was not statistically significant, addressing the root causes of rising attendance remains critical to reducing systemic strain. Potential strategies include expanding primary care access, implementing community triaging systems, and improving public health education to promote appropriate healthcare utilization. These measures could help balance patient loads and reduce delays for non-emergency procedures.

@@

## Future Work
To address the limitations of this study, future research should prioritize collecting monthly data for the years 2015 to 2019 to explore more granular trends before the anomalies introduced by COVID-19. @sec-adapted shows the potential for this further investigation with a model consisting of the yearly data from 2015 to 2019 exhibiting a much stronger $R^2$ value of 0.875. This model is small (5 observations), thus a more extensive model involving the monthly data would be beneficial. Analyzing these pre-pandemic patterns could provide a clearer understanding of the role that metrics such as beds per 1,000, physician density, and attendance rates played during a relatively stable period. Additionally, expanding the dataset to include broader healthcare metrics, such as regional funding levels, technological investments, and patient demographics, could yield a more comprehensive analysis. Future studies should also explore non-linear models or interaction effects to capture the intricate relationships between predictors. Simulation models and scenario analyses could be used to test policy interventions, such as optimizing resource allocation or improving workflow efficiency, to identify actionable strategies for reducing wait times and improving healthcare delivery.

@@

# Print the combined plot
print(combined_plot)
```
### Adapted Model {#sec-adapted}
```{r}
#| echo: false
#| output: asis
#| eval: true
#| label: tbl-modeladj
#| tbl-cap: "Linear model explaining percentage increase in wait times based on healthcare system predictors"
